Designing Qusecs Optimally: Quasi-uniform or Self-similar Layered Materials as Recursively Decomposed Solutions of Geometric Constraint Systems

### 2015-02-16 doc/GMP2015/revision_1/s00_abstract.tex
Compiles with Elsevier format. Got rid of minor spacing errors. Converted tikz figures to eps. Aligned subfigures. Etc. It's looking good! Now to consider major changes.

Crucial for the design and analysis of the abovementioned properties
is the recursive decomposition of diverse types of underlying 2D
geometric constraint systems that are \vemph{isostatic or
underconstrained (which we call collectively as \dfn{qusecs})}. While
the optimal recursive decomposition (\dfn{DR-planning}) problem is NP-
hard even for (possibly overconstrained) 2D bar-joint constraint
systems we describe an \candrpcomplexity algorithm  for qusecs. Besides
elucidating NP-hardness, the algorithm outputs \dfn{canonical} DR-
plans that meet various desirable, previously studied criteria. We
leverage recent results on Cayley configuration spaces to show that
the indecomposable systems -  that are solved at node of the optimal
DR-plan by recombining solutions to child systems - can be minimally
modified to become decomposable and have a small DR-plan, leading to
efficient realization algorithms. We show formal connections to well-
known problems such as completion of underconstrained systems. We
briefly discuss the modeling and analysis of two specific 2D layered
materials as applications of the above theory and algorithms. We
intend to make a software implementation and videos available upon
request and publicly available for the final version.
\end{abstract}

>>>>

Crucial for the design and analysis of the abovementioned properties is the recursive decomposition of diverse types of underlying 2D geometric constraint systems that are \vemph{isostatic or underconstrained (which we call collectively as \dfn{qusecs})}. While the optimal recursive decomposition (\dfn{DR-planning}) problem is NP-hard even for (possibly overconstrained) 2D bar-joint constraint systems we describe an \candrpcomplexity\ algorithm  for qusecs. Besides elucidating NP-hardness, the algorithm outputs \dfn{canonical} DR-plans that meet various desirable, previously studied criteria. We leverage recent results on Cayley configuration spaces to show that the indecomposable systems---that are solved at node of the optimal DR-plan by recombining solutions to child systems---can be minimally modified to become decomposable and have a small DR-plan, leading to efficient realization algorithms. We show formal connections to well-known problems such as completion of underconstrained systems. We briefly discuss the modeling and analysis of two specific 2D layered materials as applications of the above theory and algorithms. We intend to make a software implementation and videos available upon request and publicly available for the final version.
\end{abstract}

\begin{keyword}
    rigidity \sep
    geometric constraint solving \sep
    configuration spaces \sep
    self-similar structures \sep
    layered materials
    %
    % \PACS 71.35.-y \sep 71.35.Lk \sep 71.36.+c
\end{keyword}

% Easychair keyphrases
% canonical dr plan (380), optimal dr plan (364), constraint system (236), pinned line incidence (200), cayley configuration space (198), geometric constraint system (190), pinned line incidence graph (180), geometric constraint (166), pinned line incidence system (120), self similar (115), bar joint (113), cayley configuration (90), parent system (90), pinned line (85), strong church rosser property (80), external stress (80), rigidity matrix (80), tree decomposable graph (79), low cayley complexity (79), bar joint system (79), connected component (70), line incidence (70), orientation type (70), wikimedia common (70), cross linking (70), computational geometry (70), rigid vertex maximal proper subgraph (69), bar joint graph (63), appropriate rigidity matrix (63), plant cell wall (63)

### 2015-02-16 doc/GMP2015/revision_1/s01_introduction.tex
Compiles with Elsevier format. Got rid of minor spacing errors. Converted tikz figures to eps. Aligned subfigures. Etc. It's looking good! Now to consider major changes.

Some properties of natural and engineered materials can be analyzed by
treating them as two dimensional (2D) layers. As illustrated by the
examples below, the structure within each layer is often self-
similar~\cite{2012arXiv1204.6389G} spanning multiple scales; generally
aperiodic and quasi-uniform within any one scale; and consists of a
few repeated motifs appearing in disordered arrangements. Each layer
is not necessarily planar, i.e., it consists of multiple, inter-
constraining planar (genus 0) monolayers. Furthermore, a layer is
often  either \vemph{isostatic or underconstrained, not self-
stressed} (See Section \ref{sec:prelim} for definitions). These
properties are consistent with a self-assembled 2D structure that
minimizes mass, optimally distributes external stresses and itself
participates in the assembly of diverse and multifunctional, larger 2D
structures.

@@

\begin{figure*}\centering
% \begin{subfigure}{.31\linewidth}\centering
%   \includegraphics[width=\linewidth]{img/Axopodium_Mikrotubuli}
%   \caption{Cross section of a Heliozoa's psuedopod, formed by a spiral structure of microtubules.}
%   % https://de.wikipedia.org/wiki/Mikrotubulus#mediaviewer/File:Axopodium_Mikrotubuli.jpg
%   \label{fig:material_examples:microtubule}
% \end{subfigure}\hfill
% \begin{subfigure}{.31\linewidth}\centering
%   \includegraphics[width=\linewidth]{img/Transverse_Section_Of_Bone}
%   \caption{Cross section of an osteon (the fundamental unit of compact bone) exhibiting its hierarchical structure.}
%   % https://en.wikipedia.org/wiki/Osteon#mediaviewer/File:Transverse_Section_Of_Bone.png
%   \label{fig:material_examples:osteon}
% \end{subfigure}\hfill
\begin{subfigure}{.28\linewidth}\centering
  \includegraphics[width=\linewidth]{img/Chlamydomonas_TEM_17}
  \caption{Cross section of the Chlamydomonas algae axoneme, a cilia composed of microtubules~\cite{wikimediacommons2007cilia}.}
  \label{fig:material_examples:microtubule}
\end{subfigure}\hfill
\begin{subfigure}{.40\linewidth}\centering
  \includegraphics[width=\linewidth]{img/ligten2}
  \caption{Cross section of a tendon displaying the hierarchical structure~\cite{lecture_biosolid_mechanics}.}
  \label{fig:material_examples:tendon}
\end{subfigure}\hfill
\begin{subfigure}{.30\linewidth}\centering
  \includegraphics[width=\linewidth]{img/Rothemund-DNA-SierpinskiGasket}
  \caption{A DNA array exhibiting the Sierpinski triangle~\cite{wikimediacommons2007dna}.}
  \label{fig:material_examples:sierpinski}
\end{subfigure}
\caption{}\label{fig:material_examples}
\end{figure*}

@@

\medskip\noindent
\header{(ii) Optimal Recursive Decomposition (DR-planning).}
Recursive decomposition of geometric constraint systems has been
formalized \cite{hoffman2001decompositionI,hoffman2001decompositionII} and well-studied \cite{jermann2006decomposition,sitharam2005combinatorial} as the
\dfn {Decomposition-Recombination (DR)-planning} problem, defined formally in
Section \ref{sec:prelim}. For the above-
mentioned classes of 2D qusecs, generic rigidity is a combinatorial
property and hence each level of the decomposition should, in
principle, be achievable by a graph-based algorithm as in (1), without
involving geometric information in the constraint system. Since many
such decompositions can exist for a given constraint system, criteria
defining desirable or optimal DR-plans and DR-planning algorithms were
given in \cite{hoffman2001decompositionI}. An \dfn{optimal DR-plan} is one that minimizes
the \dfn{size} (defined formally in Section \ref{sec:prelim}, i.e.,
the maximum number of child subsystems of any parent system.
Being exponential the size, the complexity of solving the parent
constraint system from the solutions of the child systems, is overwhelmingly
dominated by it.

@@

While polynomial time algorithms were given \cite{hoffman2001decompositionI} to generate DR-
plans meeting the cluster minimality criterion, no such algorithm is
known for the latter criterion.

@@

\medskip\noindent
\header{(v) Configuration Spaces of Underconstrained Systems.}
For underconstrained 2D bar-joint and body-hyperpin qusecs obtained
from various subclasses of tree-decomposable systems, algorithms have
been developed to complete them into isostatic systems
\cite{joan-arinyo2003transforming,sitharam2005combinatorial,gao2006ctree,sitharam2010convex} and to find paths within the connected components
\cite{sitharam2011cayleyI,hidalgo2011reachability} of standard Cartesian configuration spaces.
Most of the
algorithms with formal guarantees leverage Cayley configuration space
theory \cite{sitharam2010convex,sitharam2011cayleyI,sitharam2011cayleyII} to characterize subclasses of graphs and
additional constraints that control combinatorial explosion, and
provide faithful bijective representation of connected components and
paths. These algorithms have decreasing efficiency as the subclass of
systems gets bigger, with highest efficiency for underlying partial
2-tree graphs (alternately called, tree-width 2, series-parallel, $K_4$
minor avoiding), moderate efficiency for 1 degree- of-freedom (dof)
graphs with low Cayley complexity (which include common linkages such
as the Strandbeest, Limacon and Cardioid), and decreased efficiency
for general 1-dof tree-decomposable graphs. While software suites
exist  \cite{keycurriculum1995geometer,porta2014open,siemens1999d,todd2007geometry}, no such formal algorithms and guarantees are known
for non-tree-decomposable systems.
%
\subsection{Contributions and Organization}
\label{sec:cont}

@@

  \item Also in Section \ref{sec:DRP}, we give an efficient (\candrpcomplexity) algorithm to find a canonical
  (and hence optimal) DR-plan for all 3 types of 2D qusecs mentioned
  above (Sections \ref{sec:DRP}, \ref{sec:bodypin}, and
  \ref{sec:pinnedline}). The canonical DR-plan elucidates the essence
  of the NP-hardness of finding optimal DR-plans for over-constrained
  systems. Futhermore, our optimal/canonical DR-plan satisfies desirable properties such as Cluster Minimality previously studied \cite{hoffman2001decompositionI}.

  \item In Section \ref{sec:recomb}, we give a method to deal with the
  algebraic complexity of recombining the realizations or solutions of
  child subsystems into a solution of the parent system
  \cite{sitharam2010optimized,sitharam2006well,sitharam2010reconciling}
  Specifically, we define the problem of minimally modifying the
  indecomposable recombination system so that it becomes decomposable
  via a small DR-plan and yet preserves the original solutions in an
  efficiently searchable manner. In Section \ref{sec:table}, we show formal connection to well
  known problems such as optimal completion of underconstrained
  systems
  \cite{joan-arinyo2003transforming,sitharam2005combinatorial,gao2006ctree}
  and to find paths within the connected components. When the
  modifications are bounded, we obtain new, efficient algorithms for
  realizing both isostatic and underconstrained qusecs by leveraging
  recent results about Cayley parameters in
  \cite{sitharam2010convex,sitharam2011cayleyI,sitharam2011cayleyII} (see Sections \ref{sec:2-tree-reduction} and \ref{sec:tdecomp}).

  \item In Section \ref{sec:bodypin} and \ref{sec:pinnedline}, we
  briefly describe applications of the above techniques to modeling,
  analyzing, and designing specific properties in 2D material
  layers~\cite{Jackson2008bodypin}. For Examples
  \ref{materialexample1}, \ref{materialexample2} (achieving
  isostaticity, distribution of stresses in self-similar, bar-joint
  systems); Example \ref{materialexample3} (canonical and optimal DR-
  plans for pinned line incidence systems
  \cite{sitharam2014incidence}) and Examples \ref{materialexample4},
  \ref{materialexample5} (boundary-conditions for achieving various
  desired properties of body-hyperpin systems).

@@

\noindent
\note All proofs appear in the Appendix (submitted an a permitted attachment via EasyChair).

>>>>

Some properties of natural and engineered materials can be analyzed by treating them as two dimensional (2D) layers. As illustrated by the examples below, the structure within each layer is often self-similar~\cite{2012arXiv1204.6389G} spanning multiple scales; generally aperiodic and quasi-uniform within any one scale; and consists of a few repeated motifs appearing in disordered arrangements. Each layer is not necessarily planar, i.e., it consists of multiple, inter-constraining planar (genus 0) monolayers. Furthermore, a layer is often  either \vemph{isostatic or underconstrained, not self-stressed} (See Section \ref{sec:prelim} for definitions). These properties are consistent with a self-assembled 2D structure that minimizes mass, optimally distributes external stresses and itself participates in the assembly of diverse and multifunctional, larger 2D structures.

@@

% \FigInit{}{fig:material_examples}
% \FigThreeSubfig%
%   {img/Chlamydomonas_TEM_17}
%   {Cross section of the Chlamydomonas algae axoneme, a cilia composed of microtubules~\cite{wikimediacommons2007cilia}.}
%   {fig:material_examples:microtubule}%
%   %
%   {img/ligten2}
%   {Cross section of a tendon displaying the hierarchical structure~\cite{lecture_biosolid_mechanics}.}
%   {fig:material_examples:tendon}%
%   %
%   {img/Rothemund-DNA-SierpinskiGasket}
%   {A DNA array exhibiting the Sierpinski triangle~\cite{wikimediacommons2007dna}.}
%   {fig:material_examples:sierpinski}

\ClearMyMinHeight
\SetMyMinHeight{.3}{img/Chlamydomonas_TEM_17}
\SetMyMinHeight{.3}{img/ligten2}
\SetMyMinHeight{.3}{img/Rothemund-DNA-SierpinskiGasket}

\begin{figure*}\centering%
  %
  \begin{subfigure}{0.3\linewidth}\centering
    \includegraphics[height=\myMinHeight]{img/Chlamydomonas_TEM_17}
    \caption{Cross section of the Chlamydomonas algae axoneme, a cilia composed of microtubules~\cite{wikimediacommons2007cilia}.}\label{fig:material_examples:microtubule}
  \end{subfigure}%
  %
  \hfill
  \begin{subfigure}{0.3\linewidth}\centering
    \includegraphics[height=\myMinHeight]{img/ligten2}
    \caption{Cross section of a tendon displaying the hierarchical structure~\cite{lecture_biosolid_mechanics}.}\label{fig:material_examples:tendon}
  \end{subfigure}%
  %
  \hfill
  \begin{subfigure}{0.3\linewidth}\centering
    \includegraphics[height=\myMinHeight]{img/Rothemund-DNA-SierpinskiGasket}
    \caption{A DNA array exhibiting the Sierpinski triangle~\cite{wikimediacommons2007dna}.}\label{fig:material_examples:sierpinski}
  \end{subfigure}%
  %
  \caption{}\label{fig:material_examples}
\end{figure*}%

% \FigAddSubfig%
%   {0.3}
%   {img/Chlamydomonas_TEM_17}
%   {Cross section of the Chlamydomonas algae axoneme, a cilia composed of microtubules~\cite{wikimediacommons2007cilia}.}
%   {fig:material_examples:microtubule}%
% \FigAddSubfig%
%   {0.3}
%   {img/ligten2}
%   {Cross section of a tendon displaying the hierarchical structure~\cite{lecture_biosolid_mechanics}.}
%   {fig:material_examples:tendon}%
% \FigAddSubfig%
%   {0.3}
%   {img/Rothemund-DNA-SierpinskiGasket}
%   {A DNA array exhibiting the Sierpinski triangle~\cite{wikimediacommons2007dna}.}
%   {fig:material_examples:sierpinski}
% \FigDisplay{}{fig:material_examples}

@@

\medskip\noindent
\header{(ii) Optimal Recursive Decomposition (DR-planning).}
Recursive decomposition of geometric constraint systems has been formalized \cite{hoffman2001decompositionI,hoffman2001decompositionII} and well-studied \cite{jermann2006decomposition,sitharam2005combinatorial} as the \dfn {Decomposition-Recombination (DR)-planning} problem, defined formally in Section \ref{sec:prelim}. For the abovementioned classes of 2D qusecs, generic rigidity is a combinatorial property and hence each level of the decomposition should, in principle, be achievable by a graph-based algorithm as in (1), without involving geometric information in the constraint system. Since many such decompositions can exist for a given constraint system, criteria defining desirable or optimal DR-plans and DR-planning algorithms were given in \cite{hoffman2001decompositionI}. An \dfn{optimal DR-plan} is one that minimizes the \dfn{size} (defined formally in Section \ref{sec:prelim}, i.e., the maximum number of child subsystems of any parent system. Being exponential the size, the complexity of solving the parent constraint system from the solutions of the child systems, is overwhelmingly dominated by it.

@@

While polynomial time algorithms were given \cite{hoffman2001decompositionI} to generate DR-plans meeting the cluster minimality criterion, no such algorithm is known for the latter criterion.

@@

\medskip\noindent
\header{(v) Configuration Spaces of Underconstrained Systems.}
For underconstrained 2D bar-joint and body-hyperpin qusecs obtained from various subclasses of tree-decomposable systems, algorithms have been developed to complete them into isostatic systems \cite{joan-arinyo2003transforming,sitharam2005combinatorial,gao2006ctree,sitharam2010convex} and to find paths within the connected components \cite{sitharam2011cayleyI,hidalgo2011reachability} of standard Cartesian configuration spaces. Most of the algorithms with formal guarantees leverage Cayley configuration space theory \cite{sitharam2010convex,sitharam2011cayleyI,sitharam2011cayleyII} to characterize subclasses of graphs and additional constraints that control combinatorial explosion, and provide faithful bijective representation of connected components and paths. These algorithms have decreasing efficiency as the subclass of systems gets bigger, with highest efficiency for underlying partial 2-tree graphs (alternately called, tree-width 2, series-parallel, $K_4$ minor avoiding), moderate efficiency for 1 degree-of-freedom (dof) graphs with low Cayley complexity (which include common linkages such as the Strandbeest, Limacon and Cardioid), and decreased efficiency for general 1-dof tree-decomposable graphs. While software suites exist  \cite{keycurriculum1995geometer,porta2014open,siemens1999d,todd2007geometry}, no such formal algorithms and guarantees are known for non-tree-decomposable systems.
%
\subsection{Contributions and Organization}
\label{sec:cont}

@@

  \item Also in Section \ref{sec:DRP}, we give an efficient (\candrpcomplexity) algorithm to find a canonical (and hence optimal) DR-plan for all 3 types of 2D qusecs mentioned above (Sections \ref{sec:DRP}, \ref{sec:bodypin}, and \ref{sec:pinnedline}). The canonical DR-plan elucidates the essence of the NP-hardness of finding optimal DR-plans for over-constrained systems. Furthermore, our optimal/canonical DR-plan satisfies desirable properties such as Cluster Minimality previously studied \cite{hoffman2001decompositionI}.

  \item In Section \ref{sec:recomb}, we give a method to deal with the algebraic complexity of recombining the realizations or solutions of child subsystems into a solution of the parent system \cite{sitharam2010optimized,sitharam2006well,sitharam2010reconciling}. Specifically, we define the problem of minimally modifying the indecomposable recombination system so that it becomes decomposable via a small DR-plan and yet preserves the original solutions in an efficiently searchable manner. In Section \ref{sec:table}, we show formal connection to well known problems such as optimal completion of underconstrained systems \cite{joan-arinyo2003transforming,sitharam2005combinatorial,gao2006ctree} and to find paths within the connected components. When the modifications are bounded, we obtain new, efficient algorithms for realizing both isostatic and underconstrained qusecs by leveraging recent results about Cayley parameters in \cite{sitharam2010convex,sitharam2011cayleyI,sitharam2011cayleyII} (see Sections \ref{sec:2-tree-reduction} and \ref{sec:tdecomp}).

  \item In Section \ref{sec:bodypin} and \ref{sec:pinnedline}, we briefly describe applications of the above techniques to modeling, analyzing, and designing specific properties in 2D material layers~\cite{Jackson2008bodypin}. For Examples \ref{materialexample1}, \ref{materialexample2} (achieving isostaticity, distribution of stresses in self-similar, bar-joint systems); Example \ref{materialexample3} (canonical and optimal DR-plans for pinned line incidence systems \cite{sitharam2014incidence}) and Examples \ref{materialexample4}, \ref{materialexample5} (boundary-conditions for achieving various desired properties of body-hyperpin systems).

@@

% \medskip\noindent
% \note All proofs appear in the Appendix.
% \medskip

% \noindent
% \note All proofs appear in the Appendix.

\sidenote{All proofs appear in \ref{sec:appendix:a}.}

### 2015-02-16 doc/GMP2015/revision_1/s02_preliminaries.tex
Compiles with Elsevier format. Got rid of minor spacing errors. Converted tikz figures to eps. Aligned subfigures. Etc. It's looking good! Now to consider major changes.

\noindent
\note Geometric constraint systems can also have
inequalities in addition to equations, where the parameters in $\delta$ are small intervals of
values rather than exact values.

@@

One important property of generic constraint system or (hyper)graph
%that depends on the appropriate rigidity matrix of indeterminates
is \dfn{rigidity} (\note we refer to these as properties of
the constraint system or as properties of the underlying (hyper)graph
interchangeably), i.e., the realizations or solutions of the
corresponding constraint system being generically isolated and zero-
dimensional.%(if the constraint system has equalities and exact values for the parameters $\delta$).
The result by Asimow and Roth \cite{asimow1978rigidity}  shows a constraint
(hyper)graph is rigid if it is generically \dfn{infinitesimally
rigid}, i.e.\ the number of independent rows of its appropriate
rigidity matrix is at least the number of columns less the number of
rigid body motions, which is 3 for bar-joint systems.
%This condition is called generic {\em infinitesimal rigidity}.

\noindent
\note If the constraint system has inequalities, i.e, if the $\delta$
lie in an interval, the definition of rigidity is approximate: the
solutions are isolated small, full-dimensional  connected components.
% generic rigidity, i.e.\ there exist at most finitely many solutions to the system at generic points,  %to the algebraic system

@@

Defining the combinatorial independence of a subset of edges $E'
\subseteq E$ to be the generic independence of corresponding rows in
the appropriate rigidity matrix of indeterminates, we obtain the
\dfn{rigidity matroid} of a constraint (hyper)graph $G = (V,E)$.%The 2-dimensional {\em rigidity matroid} of a constraint (hyper)graph $G = (V,E)$ is a linear matroid  on $E$,
%where a subset of edges $E' \subseteq E$ is {\em independent} in the matroid,
%if the set of corresponding rows in the appropriate rigidity matrix of indeterminates are linearly independent.
There are various results on combinatorial characterization of
independence and rigidity  and rigidity matroids for different types
of graphs. For bar-joint rigidity matroid, the famous Laman's theorem
\cite{laman1970graphs} states that the underlying graph is isostatic if and only if
$|E| = 2|V|-3$ and $|E'| \le 2|V'|-3$ for every induced subgraph with
at least 2 vertices. The result by Lovasz and Yemini \cite{lovasz1982generic}  shows
that all 6-vertex-connected graphs are rigid in the plane. For bar-
body rigidity matroid, Tay \cite{tay1976rigidity} proved that the underlying
multigraph is isostatic if and only if it can be decomposed as $3$
edge disjoint spanning trees. White and Whiteley \cite{white1987algebraic} gave the
same characterization using a different technique to study the
algebraic-geometric conditions of genericity, called pure condition.
Lee, Streinu and Theran \cite{lee2007graded} defined the \dfn{$(k,l)$-sparsity
matroid}, where a hypergraph $G$ is called \dfn{$(k,l)$-sparse} if
$|E'| \le k|V'| - l$ for any induced subgraph $(V',E')$ with at least
2 vertices, and \dfn{$(k,l)$-tight} if is $(k,l)$-sparse and $|E| =
k|V| - l$. In general, given a $d$-uniform hypergraph, a
$(k,l)$-sparsity condition is matroidal as long as $l \le dk-1$.

@@

\medskip\noindent
\note In the remainder of this section and the next three
sections~\ref{sec:DRP} and \ref{sec:recomb} we only consider bar-joint
qusecs and graphs. Relevant formal analogies for the other 2 types of
qusecs and (hyper)graphs are given in the subsequent 2 sections on
Applications.

@@

\subsection{Decomposition-Recombination (DR) Plans}
%
%$G=(V,E,w_V,w_E)$ is a set of vertices, $V$, and a set of edges, $E$, defined as a tuple of vertices from $V$ (undirected). Additionally, there are two weight functions, $w_V: V \to \mathbb{R}^+$ and $w_E: E \to \mathbb{R}^+$. The \textbf{density} of graph $G$ is $d(G) = \sum_{e\in E}{w_E(e)} - \sum_{v\in V}{w_V(v)}$.
%
%Given a constant $k$, a graph $G$ is:
%\begin{itemize}
%    \item \textbf{independent} (or \textbf{sparse}) if, for all non-trivial subgraphs $S\subseteq G$, $d(S) \leq k$.
%
%    \item \textbf{overconstrained} if it is not independent.
%
%    \item \textbf{wellconstrained} (or \textbf{tight}) if $G$ is independent and $d(G)=k$.
%
%    \item \textbf{rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is wellconstrained.
%
%    \item \textbf{underconstrained} if $G$ is independent and not rigid.
%
%    % \item \textbf{trivial} if (1) it is overconstrained graph and (2) all of its subgraphs are also trivial. Trivial graphs are input.
%\end{itemize}
%
% \begin{definition}
%     A graph is \textbf{overconstrained} if, given constant $k$, there exists some
%     % induced
%     subgraph $S\subseteq G$ such that $d(S) > k$.
% \end{definition}
%
% \begin{definition}
%     A graph is \textbf{wellconstrained} if, given constant $k$, $d(G) = k$ and for all
%     % induced
%     subgraphs $S\subseteq G$, (1) $d(S)\leq k$, or (2) given a set of trivial graphs $T$, $S$ is isomorphic to one of the graphs in $T$ (i.e.\ $S$ is trivial).
% \end{definition}
%
%
%
% \begin{definition}
%     A \textbf{trivial} graph is ill-defined in the general case. The only strict requirements are: (1) it must be an overconstrained graph and (2) all of its subgraphs are also trivial.
%     \todo{Maybe leave the following out until it's needed later?}
%     In the familiar geometric cases of $d$-dimensional space, all vertex weights are $d$, all edge weights are $1$, and constant $k= -{{d+1}\choose{2}}$. Trivial graphs for 2D would be a single vertex. Trivial graphs in 3D would be a vertex and an edge (2 vertices with an edge between). Etc. These trivial graphs capture the notion of the rotational symmetry that exists in geometric spaces.
% \end{definition}
%
\begin{definition}\label{def:drp}
    The \dfn{decomposition-recombination \mbox{(DR-)} plan} of graph
    $G$, \cite{hoffman2001decompositionI} $\drp{G}$ (see Figure \ref{fig:c2c3ofk33s}, is defined as a forest that has the
    following properties:
    \begin{enumerate}
        \item Each node represents/contains/is a rigid vertex-induced
        subgraph of $G$.
        % \item For a node, $C$, that is  a non-trivial graph, its $N$
        % children, $C_1, \ldots, C_N$, are rigid vertex-maximal proper
        % subgraphs of $C$.
        \item Node $C$ has children $C_1,\ldots,C_N$ such that $\bigcup_{i=1}^N{C_i}=C$.
        \item A leaf node is a single vertex.
A \dfn{trivial} graph is empty or a single vertex.
        \item A root node is a maximal rigid subgraph of $G$.
    \end{enumerate}
%
    % It can also be described recursively as: the root is $G$, its children are the trivial subgraphs and the DR-plans of its wellconstrained vertex-maximal proper subgraphs whose union is $G$ itself.
%
    A DR-plan is \dfn{complete} if it satisfies an additional rule:
        \item For a node $C$ that is  a non-trivial graph, its $N$
        children, $C_1, \ldots, C_N$, are all of the rigid vertex-
        maximal proper subgraphs of $C$. This makes Rule 2 implicit,
        and in fact, $C$ contains all the edges in the union of its
        children as well.
        We denote a complete DR-plan of $G$ as $\comdrp{G}$
    A DR-plan is \dfn{optimal} if it minimizes the maximum fan-in over all nodes in the tree.
    The maximum  fan-in is called the \dfn{size} of the DR-plan.
    We denote an optimal DR-plan of $G$ as $\optdrp{G}$.

@@

\noindent
\note In Section \ref{sec:DRP}, $G$ is assumed to be isostatic.

>>>>

% \noindent
% \note Geometric constraint systems can also have inequalities in addition to equations, where the parameters in $\delta$ are small intervals of values rather than exact values.

\sidenote{Geometric constraint systems can also have inequalities in addition to equations, where the parameters in $\delta$ are small intervals of values rather than exact values.}

@@

One important property of generic constraint system or (hyper)graph
%that depends on the appropriate rigidity matrix of indeterminates
is \dfn{rigidity} (\note we refer to these as properties of the constraint system or as properties of the underlying (hyper)graph interchangeably), i.e., the realizations or solutions of the corresponding constraint system being generically isolated and zero-dimensional.%(if the constraint system has equalities and exact values for the parameters $\delta$).
The result by Asimow and Roth \cite{asimow1978rigidity}  shows a constraint (hyper)graph is rigid if it is generically \dfn{infinitesimally rigid}, i.e.\ the number of independent rows of its appropriate rigidity matrix is at least the number of columns less the number of rigid body motions, which is 3 for bar-joint systems.
%This condition is called generic {\em infinitesimal rigidity}.

% \noindent
% \note If the constraint system has inequalities, i.e, if the $\delta$ lie in an interval, the definition of rigidity is approximate: the solutions are isolated small, full-dimensional  connected components.

\sidenote{If the constraint system has inequalities, i.e, if the $\delta$ lie in an interval, the definition of rigidity is approximate: the solutions are isolated small, full-dimensional  connected components.}
% generic rigidity, i.e.\ there exist at most finitely many solutions to the system at generic points,  %to the algebraic system

@@

Defining the combinatorial independence of a subset of edges $E'\subseteq E$ to be the generic independence of corresponding rows in the appropriate rigidity matrix of indeterminates, we obtain the \dfn{rigidity matroid} of a constraint (hyper)graph $G = (V,E)$.%The 2-dimensional {\em rigidity matroid} of a constraint (hyper)graph $G = (V,E)$ is a linear matroid  on $E$,
%where a subset of edges $E' \subseteq E$ is {\em independent} in the matroid,
%if the set of corresponding rows in the appropriate rigidity matrix of indeterminates are linearly independent.
There are various results on combinatorial characterization of independence and rigidity  and rigidity matroids for different types of graphs. For bar-joint rigidity matroid, the famous Laman's theorem \cite{laman1970graphs} states that the underlying graph is isostatic if and only if $|E| = 2|V|-3$ and $|E'| \le 2|V'|-3$ for every induced subgraph with at least 2 vertices. The result by Lovasz and Yemini \cite{lovasz1982generic}  shows that all 6-vertex-connected graphs are rigid in the plane. For bar-body rigidity matroid, Tay \cite{tay1976rigidity} proved that the underlying multigraph is isostatic if and only if it can be decomposed as $3$ edge disjoint spanning trees. White and Whiteley \cite{white1987algebraic} gave the same characterization using a different technique to study the algebraic-geometric conditions of genericity, called pure condition. Lee, Streinu and Theran \cite{lee2007graded} defined the \dfn{$(k,l)$-sparsity matroid}, where a hypergraph $G$ is called \dfn{$(k,l)$-sparse} if $|E'| \le k|V'| - l$ for any induced subgraph $(V',E')$ with at least 2 vertices, and \dfn{$(k,l)$-tight} if is $(k,l)$-sparse and $|E| = k|V| - l$. In general, given a $d$-uniform hypergraph, a $(k,l)$-sparsity condition is matroidal as long as $l \le dk-1$.

@@

% \medskip\noindent
% \note In the remainder of this section and sections~\ref{sec:DRP} and \ref{sec:recomb} we only consider bar-joint qusecs and graphs. Relevant formal analogies for the other 2 types of qusecs and (hyper)graphs are given in the subsequent 2 sections on Applications.

\sidenote{In the remainder of this section and sections~\ref{sec:DRP} and \ref{sec:recomb} we only consider bar-joint qusecs and graphs. Relevant formal analogies for the other 2 types of qusecs and (hyper)graphs are given in the subsequent 2 sections on Applications.}

@@

\subsection{Decomposition-Recombination (DR) Plans}
%
%$G=(V,E,w_V,w_E)$ is a set of vertices, $V$, and a set of edges, $E$, defined as a tuple of vertices from $V$ (undirected). Additionally, there are two weight functions, $w_V: V \to \mathbb{R}^+$ and $w_E: E \to \mathbb{R}^+$. The \textbf{density} of graph $G$ is $d(G) = \sum_{e\in E}{w_E(e)} - \sum_{v\in V}{w_V(v)}$.
%
%Given a constant $k$, a graph $G$ is:
%\begin{itemize}
%    \item \textbf{independent} (or \textbf{sparse}) if, for all non-trivial subgraphs $S\subseteq G$, $d(S) \leq k$.
%
%    \item \textbf{overconstrained} if it is not independent.
%
%    \item \textbf{wellconstrained} (or \textbf{tight}) if $G$ is independent and $d(G)=k$.
%
%    \item \textbf{rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is wellconstrained.
%
%    \item \textbf{underconstrained} if $G$ is independent and not rigid.
%
%    % \item \textbf{trivial} if (1) it is overconstrained graph and (2) all of its subgraphs are also trivial. Trivial graphs are input.
%\end{itemize}
%
% \begin{definition}
%     A graph is \textbf{overconstrained} if, given constant $k$, there exists some
%     % induced
%     subgraph $S\subseteq G$ such that $d(S) > k$.
% \end{definition}
%
% \begin{definition}
%     A graph is \textbf{wellconstrained} if, given constant $k$, $d(G) = k$ and for all
%     % induced
%     subgraphs $S\subseteq G$, (1) $d(S)\leq k$, or (2) given a set of trivial graphs $T$, $S$ is isomorphic to one of the graphs in $T$ (i.e.\ $S$ is trivial).
% \end{definition}
%
%
%
% \begin{definition}
%     A \textbf{trivial} graph is ill-defined in the general case. The only strict requirements are: (1) it must be an overconstrained graph and (2) all of its subgraphs are also trivial.
%     \todo{Maybe leave the following out until it's needed later?}
%     In the familiar geometric cases of $d$-dimensional space, all vertex weights are $d$, all edge weights are $1$, and constant $k= -{{d+1}\choose{2}}$. Trivial graphs for 2D would be a single vertex. Trivial graphs in 3D would be a vertex and an edge (2 vertices with an edge between). Etc. These trivial graphs capture the notion of the rotational symmetry that exists in geometric spaces.
% \end{definition}
%
\begin{definition}\label{def:drp}
    The \dfn{decomposition-recombination \mbox{(DR-)} plan} of graph $G$, \cite{hoffman2001decompositionI} $\drp{G}$ (see Figure \ref{fig:c2c3ofk33s}, is defined as a forest that has the following properties:
    \begin{enumerate}
        \item Each node represents/contains/is a rigid vertex-induced subgraph of $G$.
        % \item For a node, $C$, that is  a non-trivial graph, its $N$
        % children, $C_1, \ldots, C_N$, are rigid vertex-maximal proper
        % subgraphs of $C$.
        \item Node $C$ has children $C_1,\ldots,C_N$ such that $\bigcup_{i=1}^N{C_i}=C$.
        \item A leaf node is a single vertex. A \dfn{trivial} graph is empty or a single vertex.
        \item A root node is a maximal rigid subgraph of $G$.
    \end{enumerate}
%
    % It can also be described recursively as: the root is $G$, its children are the trivial subgraphs and the DR-plans of its wellconstrained vertex-maximal proper subgraphs whose union is $G$ itself.
%
    A DR-plan is \dfn{complete} if it satisfies an additional rule:
        \item For a node $C$ that is  a non-trivial graph, its $N$ children, $C_1, \ldots, C_N$, are all of the rigid vertex-maximal proper subgraphs of $C$. This makes Rule 2 implicit, and in fact, $C$ contains all the edges in the union of its children as well. We denote a complete DR-plan of $G$ as $\comdrp{G}$
    A DR-plan is \dfn{optimal} if it minimizes the maximum fan-in over all nodes in the tree. The maximum  fan-in is called the \dfn{size} of the DR-plan. We denote an optimal DR-plan of $G$ as $\optdrp{G}$.

@@

\sidenote{In Section \ref{sec:DRP}, $G$ is assumed to be isostatic.}
